---
title: "Validation of 2022 CPS Turnout Estimates Against McDonald VEP Figures"
author: "Paul Gronke and Frank Adonteng"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(cpsvote)
library(dplyr)
library(srvyr)
library(googlesheets4)
library(knitr)
library(kableExtra)
library(ggplot2)
```

# Overview

This document compares state-level voter turnout estimates from the 2022 CPS
Voting and Registration Supplement — using the **standard Census survey weight**
(`WEIGHT`) — against the authoritative Voter Eligible Population (VEP) turnout
figures from Dr. Michael McDonald's United States Elections Project.

Using the standard weight rather than the Hur-Achen reweighted weight allows
us to observe the raw discrepancy between what the CPS reports and the true
turnout figures. Because the CPS is known to overreport turnout (due to social
desirability bias and differential non-response), we expect CPS estimates to
be systematically higher than McDonald's VEP figures in most states. The size
of this overestimation varies by state and is precisely what the Hur-Achen
reweighting procedure corrects for.

The turnout variable used is `hurachen_turnout`, which drops respondents who
answered Don't Know, Refused, or No Response (treating them as missing rather
than as non-voters). This isolates the effect of the **weight** on the
discrepancy, holding the coding of the turnout variable constant.

# CPS 2022 Turnout Estimates (Standard Weight)

```{r load-cps, cache=TRUE}
cps_2022 <- cps_load_basic(years = 2022) %>%
  as_survey_design(weights = WEIGHT)   # standard Census weight, NOT turnout_weight

cps_turnout <- cps_2022 %>%
  group_by(STATE) %>%
  summarize(
    cps_turnout_pct = survey_mean(hurachen_turnout == "YES", na.rm = TRUE)
  ) %>%
  select(STATE, cps_turnout_pct) %>%
  mutate(cps_turnout_pct = round(cps_turnout_pct * 100, 1))
```

# McDonald VEP Turnout Estimates

```{r load-mcdonald, cache=TRUE}
gs4_deauth()

gid_2022 <- "17iSZIBPP6jj0C2wcqpym9wNuNCtTiqM_tMweMjmvces"

mcdonald_2022 <- read_sheet(
  gid_2022,
  range = "A3:N54",
  col_names = c(
    "state_name", "highestoffice", "status", "results_source",
    "pct_highestoffice_vep", "vep", "vap", "pct_noncitizen",
    "prison", "probation", "parole", "total_ineligible_felon",
    "overseas_eligible", "state_abb"
  )
) %>%
  filter(state_abb != "US") %>%
  transmute(
    STATE           = state_abb,
    state_name      = stringr::str_remove_all(state_name, "\\*$"),
    mcd_turnout_pct = round(as.numeric(pct_highestoffice_vep) * 100, 1)
  )
```

# State-by-State Comparison

```{r join}
comparison <- left_join(mcdonald_2022, cps_turnout, by = "STATE") %>%
  mutate(difference = cps_turnout_pct - mcd_turnout_pct) %>%
  arrange(state_name)
```

The table below shows, for each state:

- **McDonald VEP (%)** — authoritative VEP turnout from the Elections Project
- **CPS Standard Weight (%)** — CPS estimate using the standard Census weight
- **Difference (pp)** — CPS minus McDonald (positive = CPS overestimates)

Differences greater than 5 percentage points are highlighted in red.

```{r table}
comparison %>%
  select(
    State                    = state_name,
    `McDonald VEP (%)`       = mcd_turnout_pct,
    `CPS Std. Weight (%)`    = cps_turnout_pct,
    `Difference (pp)`        = difference
  ) %>%
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    align     = c("l", "r", "r", "r"),
    caption   = "2022 State Voter Turnout: CPS Standard Weight vs. McDonald VEP Estimates"
  ) %>%
  kable_styling(
    latex_options = c("repeat_header", "striped"),
    font_size     = 9
  ) %>%
  column_spec(4, bold = TRUE,
              color = ifelse(abs(comparison$difference) > 5, "red", "black"))
```

\newpage

# Summary Statistics

```{r summary}
summary_stats <- comparison %>%
  summarize(
    `Mean difference (pp)`   = round(mean(difference, na.rm = TRUE), 2),
    `Median difference (pp)` = round(median(difference, na.rm = TRUE), 2),
    `SD of difference (pp)`  = round(sd(difference, na.rm = TRUE), 2),
    `Max overestimate (pp)`  = round(max(difference, na.rm = TRUE), 2),
    `Min difference (pp)`    = round(min(difference, na.rm = TRUE), 2),
    `States overestimating`  = sum(difference > 0, na.rm = TRUE),
    `States within 5pp`      = sum(abs(difference) <= 5, na.rm = TRUE),
    `States > 5pp off`       = sum(abs(difference) > 5, na.rm = TRUE)
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "Statistic", values_to = "Value")

kable(summary_stats, booktabs = TRUE,
      caption = "Summary of CPS (standard weight) vs. McDonald Differences (percentage points)") %>%
  kable_styling(latex_options = "hold_position")
```

A positive mean difference confirms the well-documented CPS overreporting
tendency. The mean and median discrepancy reflect the degree of bias that the
Hur-Achen reweighting procedure corrects for. **Min difference** is the state
where CPS overestimates the least (or, if negative, underestimates the most);
a positive min difference means no state actually falls below McDonald's VEP
figure.

```{r scatter, fig.height=5, fig.cap="CPS (standard weight) vs. McDonald VEP turnout by state, 2022. The dashed line is the 45-degree line of perfect agreement. Points above the line indicate CPS overestimation. States labeled in red differ by more than 5 percentage points."}
comparison %>%
  mutate(label = ifelse(abs(difference) > 15, STATE, NA)) %>%
  ggplot(aes(x = mcd_turnout_pct, y = cps_turnout_pct)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = abs(difference) > 5), size = 2) +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, color = "red",
                           na.rm = TRUE, max.overlaps = 20) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "red"),
                     guide = "none") +
  scale_x_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 70)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 80)) +
  labs(
    x        = "McDonald VEP Turnout (%)",
    y        = "CPS Standard Weight Turnout (%)",
    title    = "CPS vs. McDonald Turnout Estimates, 2022",
    subtitle = "Using standard Census survey weight; points above the line = CPS overestimates"
  ) +
  theme_bw()
```

# Largest Discrepancies

States differing by more than 5 percentage points are listed below, sorted by
absolute discrepancy. These are the states where non-response bias or
overreporting is most pronounced, and where the Hur-Achen reweighting makes
the largest practical difference.

```{r flagged}
flagged <- comparison %>%
  filter(abs(difference) > 5) %>%
  arrange(desc(abs(difference))) %>%
  select(
    State                 = state_name,
    `McDonald VEP (%)`    = mcd_turnout_pct,
    `CPS Std. Weight (%)` = cps_turnout_pct,
    `Difference (pp)`     = difference
  )

if (nrow(flagged) == 0) {
  cat("No states differ by more than 5 percentage points.")
} else {
  kable(flagged, booktabs = TRUE,
        caption = "States with discrepancy greater than 5 percentage points") %>%
    kable_styling(latex_options = "hold_position")
}
```

# Notes

- The standard Census weight (`WEIGHT`) is used here, **not** the Hur-Achen
  reweighted weight (`turnout_weight`). The reweighted weight is specifically
  designed to eliminate these discrepancies by post-stratifying to McDonald's
  VEP figures; using it here would produce near-perfect agreement by
  construction and would not illuminate the underlying bias.
- The `hurachen_turnout` coding is used (DK/Refused/No Response treated as
  missing). To see the additional effect of the Census's own coding (which
  treats these as No), substitute `cps_turnout` for `hurachen_turnout`.
- McDonald's VEP data for 2022 retrieved from:
  `https://docs.google.com/spreadsheets/d/17iSZIBPP6jj0C2wcqpym9wNuNCtTiqM_tMweMjmvces`
