---
title: "Validation of 2022 CPS Turnout Estimates: Two Comparisons"
author: "Paul Gronke and Frank Adonteng"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(cpsvote)
library(dplyr)
library(srvyr)
library(googlesheets4)
library(knitr)
library(kableExtra)
library(ggplot2)
```

# Overview

This document performs two validation checks on the 2022 `cpsvote` estimates.

**Comparison 1 — McDonald VEP (Sections 2–6):** Compares CPS estimates using
the standard Census weight (`WEIGHT`) against Dr. Michael McDonald's VEP
turnout figures. Because the CPS overreports turnout, we expect CPS estimates
to be systematically higher. The size of this gap is precisely what the
Hur-Achen reweighting corrects.

**Comparison 2 — Census Table 4a (Section 7):** Compares CPS estimates using
the standard Census weight and the Census's own turnout coding (`cps_turnout`,
which treats DK/Refused/No Response as "No") against the Census Bureau's
officially published Table 4a figures for 2022. These should match precisely.
A discrepancy here indicates a bug in the package.

The turnout variable used is `hurachen_turnout`, which drops respondents who
answered Don't Know, Refused, or No Response (treating them as missing rather
than as non-voters). This isolates the effect of the **weight** on the
discrepancy, holding the coding of the turnout variable constant.

# CPS 2022 Turnout Estimates (Standard Weight)

```{r load-cps, cache=TRUE}
cps_2022 <- cps_load_basic(years = 2022) %>%
  as_survey_design(weights = WEIGHT)   # standard Census weight, NOT turnout_weight

cps_turnout <- cps_2022 %>%
  group_by(STATE) %>%
  summarize(
    cps_turnout_pct = survey_mean(hurachen_turnout == "YES", na.rm = TRUE)
  ) %>%
  select(STATE, cps_turnout_pct) %>%
  mutate(cps_turnout_pct = round(cps_turnout_pct * 100, 1))
```

# McDonald VEP Turnout Estimates

```{r load-mcdonald, cache=TRUE}
gs4_deauth()

gid_2022 <- "17iSZIBPP6jj0C2wcqpym9wNuNCtTiqM_tMweMjmvces"

mcdonald_2022 <- read_sheet(
  gid_2022,
  range = "A3:N54",
  col_names = c(
    "state_name", "highestoffice", "status", "results_source",
    "pct_highestoffice_vep", "vep", "vap", "pct_noncitizen",
    "prison", "probation", "parole", "total_ineligible_felon",
    "overseas_eligible", "state_abb"
  )
) %>%
  filter(state_abb != "US") %>%
  transmute(
    STATE           = state_abb,
    state_name      = stringr::str_remove_all(state_name, "\\*$"),
    mcd_turnout_pct = round(as.numeric(pct_highestoffice_vep) * 100, 1)
  )
```

# State-by-State Comparison

```{r join}
comparison <- left_join(mcdonald_2022, cps_turnout, by = "STATE") %>%
  mutate(difference = cps_turnout_pct - mcd_turnout_pct) %>%
  arrange(state_name)
```

The table below shows, for each state:

- **McDonald VEP (%)** — authoritative VEP turnout from the Elections Project
- **CPS Standard Weight (%)** — CPS estimate using the standard Census weight
- **Difference (pp)** — CPS minus McDonald (positive = CPS overestimates)

Differences greater than 5 percentage points are highlighted in red.

```{r table}
comparison %>%
  select(
    State                    = state_name,
    `McDonald VEP (%)`       = mcd_turnout_pct,
    `CPS Std. Weight (%)`    = cps_turnout_pct,
    `Difference (pp)`        = difference
  ) %>%
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    align     = c("l", "r", "r", "r"),
    caption   = "2022 State Voter Turnout: CPS Standard Weight vs. McDonald VEP Estimates"
  ) %>%
  kable_styling(
    latex_options = c("repeat_header", "striped"),
    font_size     = 9
  ) %>%
  column_spec(4, bold = TRUE,
              color = ifelse(abs(comparison$difference) > 5, "red", "black"))
```

\newpage

# Summary Statistics

```{r summary}
summary_stats <- comparison %>%
  summarize(
    `Mean difference (pp)`   = round(mean(difference, na.rm = TRUE), 2),
    `Median difference (pp)` = round(median(difference, na.rm = TRUE), 2),
    `SD of difference (pp)`  = round(sd(difference, na.rm = TRUE), 2),
    `Max overestimate (pp)`  = round(max(difference, na.rm = TRUE), 2),
    `Min difference (pp)`    = round(min(difference, na.rm = TRUE), 2),
    `States overestimating`  = sum(difference > 0, na.rm = TRUE),
    `States within 5pp`      = sum(abs(difference) <= 5, na.rm = TRUE),
    `States > 5pp off`       = sum(abs(difference) > 5, na.rm = TRUE)
  ) %>%
  tidyr::pivot_longer(everything(), names_to = "Statistic", values_to = "Value")

kable(summary_stats, booktabs = TRUE,
      caption = "Summary of CPS (standard weight) vs. McDonald Differences (percentage points)") %>%
  kable_styling(latex_options = "hold_position")
```

A positive mean difference confirms the well-documented CPS overreporting
tendency. The mean and median discrepancy reflect the degree of bias that the
Hur-Achen reweighting procedure corrects for. **Min difference** is the state
where CPS overestimates the least (or, if negative, underestimates the most);
a positive min difference means no state actually falls below McDonald's VEP
figure.

```{r scatter, fig.height=5, fig.cap="CPS (standard weight) vs. McDonald VEP turnout by state, 2022. The dashed line is the 45-degree line of perfect agreement. Points above the line indicate CPS overestimation. States labeled in red differ by more than 5 percentage points."}
comparison %>%
  mutate(label = ifelse(abs(difference) > 15, STATE, NA)) %>%
  ggplot(aes(x = mcd_turnout_pct, y = cps_turnout_pct)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = abs(difference) > 5), size = 2) +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, color = "red",
                           na.rm = TRUE, max.overlaps = 20) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "red"),
                     guide = "none") +
  scale_x_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 70)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 80)) +
  labs(
    x        = "McDonald VEP Turnout (%)",
    y        = "CPS Standard Weight Turnout (%)",
    title    = "CPS vs. McDonald Turnout Estimates, 2022",
    subtitle = "Using standard Census survey weight; points above the line = CPS overestimates"
  ) +
  theme_bw()
```

Figure 1 illustrates the well-known CPS overreporting bias: virtually every state falls above the 45-degree line, confirming that the standard Census weight systematically overstates turnout relative to McDonald's VEP benchmarks. This gap arises because non-voters are more likely to refuse CPS participation, and the standard weight does not correct for this differential non-response. Figure 3 below reruns the same comparison using the Hur-Achen reweighted weight (`turnout_weight`), which post-stratifies respondents to match VEP figures. The shift of points toward the 45-degree line demonstrates the core value of the `cpsvote` package: providing a corrected weight that brings CPS-based state turnout estimates into close alignment with authoritative benchmarks.

```{r ha-cps, cache=TRUE}
cps_ha <- cps_load_basic(years = 2022) %>%
  as_survey_design(weights = turnout_weight)

cps_ha_turnout <- cps_ha %>%
  group_by(STATE) %>%
  summarize(
    ha_turnout_pct = survey_mean(hurachen_turnout == "YES", na.rm = TRUE)
  ) %>%
  select(STATE, ha_turnout_pct) %>%
  mutate(ha_turnout_pct = round(ha_turnout_pct * 100, 1))

comparison_ha <- left_join(mcdonald_2022, cps_ha_turnout, by = "STATE") %>%
  mutate(difference_ha = ha_turnout_pct - mcd_turnout_pct)
```

```{r ha-scatter, fig.height=5, fig.cap="CPS (Hur-Achen weight) vs. McDonald VEP turnout by state, 2022. The dashed line is the 45-degree line of perfect agreement. The Hur-Achen reweighting pulls points markedly closer to the line compared with Figure 1. States labeled in red still differ by more than 5 percentage points after reweighting."}
comparison_ha %>%
  mutate(label = ifelse(abs(difference_ha) > 5, STATE, NA)) %>%
  ggplot(aes(x = mcd_turnout_pct, y = ha_turnout_pct)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = abs(difference_ha) > 5), size = 2) +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, color = "red",
                           na.rm = TRUE, max.overlaps = 20) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "red"),
                     guide = "none") +
  scale_x_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 70)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(20, 80)) +
  labs(
    x        = "McDonald VEP Turnout (%)",
    y        = "CPS Hur-Achen Weight Turnout (%)",
    title    = "CPS (Hur-Achen weight) vs. McDonald Turnout Estimates, 2022",
    subtitle = "Using Hur-Achen reweighted survey weight; points near line = close agreement"
  ) +
  theme_bw()
```

# Largest Discrepancies

States differing by more than 5 percentage points are listed below, sorted by
absolute discrepancy. These are the states where non-response bias or
overreporting is most pronounced, and where the Hur-Achen reweighting makes
the largest practical difference.

```{r flagged}
flagged <- comparison %>%
  filter(abs(difference) > 5) %>%
  arrange(desc(abs(difference))) %>%
  select(
    State                 = state_name,
    `McDonald VEP (%)`    = mcd_turnout_pct,
    `CPS Std. Weight (%)` = cps_turnout_pct,
    `Difference (pp)`     = difference
  )

if (nrow(flagged) == 0) {
  cat("No states differ by more than 5 percentage points.")
} else {
  kable(flagged, booktabs = TRUE,
        caption = "States with discrepancy greater than 5 percentage points") %>%
    kable_styling(latex_options = "hold_position")
}
```

# Comparison with Official Census Bureau Estimates (Table 4a)

This is a **consistency check**. If `cpsvote` correctly reproduces the Census
Bureau's own methodology — standard `WEIGHT` plus Census coding of
DK/Refused/No Response as "No" (`cps_turnout`) — then our state-level
estimates should match Table 4a (Reported Voting and Registration of the Total
Voting-Age Population, for States: November 2022) precisely. Any meaningful
discrepancy indicates a problem in the package implementation.

Note: `hurachen_turnout` (used in Sections 2–6) treats DK/Refused/No Response
as missing; `cps_turnout` (used here) treats them as "No", exactly as the
Census does in its published reports.

```{r census-data, cache=TRUE}
census_url  <- "https://www2.census.gov/programs-surveys/cps/tables/p20/586/vote04a_2022.xlsx"
census_file <- tempfile(fileext = ".xlsx")
download.file(census_url, census_file, quiet = TRUE, mode = "wb")

census_raw <- readxl::read_excel(census_file, col_names = FALSE, skip = 4)

state_lookup <- tibble::tibble(
  state_name = c(toupper(state.name), "DISTRICT OF COLUMBIA"),
  state_abb  = c(state.abb, "DC")
)

# Column ...12 = "Voted - Percent of citizen population"
census_table4a <- census_raw %>%
  select(state_name = `...1`, voted_pct_citizen = `...12`) %>%
  filter(!is.na(state_name), state_name %in% state_lookup$state_name) %>%
  left_join(state_lookup, by = "state_name") %>%
  mutate(census_pct = as.numeric(voted_pct_citizen)) %>%
  select(STATE = state_abb, state_name, census_pct)
```

```{r census-cps-estimates}
# Unadjusted estimates (raw WEIGHT, Census coding) — kept for reference only
our_census_est_raw <- cps_2022 %>%
  group_by(STATE) %>%
  summarize(our_pct_raw = survey_mean(cps_turnout == "YES", na.rm = TRUE)) %>%
  mutate(STATE       = as.character(STATE),
         our_pct_raw = round(our_pct_raw * 100, 1))
```

```{r post-stratify}
# WHY postStratify() CANNOT FIX TURNOUT PROPORTIONS:
# postStratify() multiplies every weight in a stratum by the same constant k.
# For a proportion p = sum(w * x) / sum(w), this gives:
#   p_new = sum(k*w * x) / sum(k*w) = p   [k cancels]
# Proportions are therefore unchanged regardless of k.
#
# THE CORRECT FIX matches how Census Table 4a is computed:
#   Numerator   = sum(WEIGHT for YES voters)       [from CPS survey]
#   Denominator = Census citizen VAP (Table 4a col 3)  [external control total]
# We substitute the external denominator directly.

# --- Step 1: Census citizen VAP by state (in persons) ---
citizen_vap_targets <- census_raw %>%
  select(state_name = `...1`, citizen_pop_thou = `...3`) %>%
  filter(!is.na(state_name), state_name %in% state_lookup$state_name) %>%
  left_join(state_lookup, by = "state_name") %>%
  transmute(STATE            = state_abb,
            census_vap_persons = as.numeric(citizen_pop_thou) * 1000)

# --- Step 2: VRS respondents (WEIGHT already scaled by cps_label: divided by 10,000) ---
vrs_raw <- cps_2022$variables %>%
  filter(!is.na(cps_turnout)) %>%
  mutate(STATE = as.character(STATE))

# --- Step 3: BEFORE — sum of WEIGHT for VRS respondents by state ---
denom_before <- vrs_raw %>%
  group_by(STATE) %>%
  summarize(before_persons = round(sum(WEIGHT)), .groups = "drop")

# Verification table: BEFORE vs AFTER (Census citizen VAP)
denom_verify <- left_join(denom_before, citizen_vap_targets, by = "STATE") %>%
  mutate(shortfall_k   = round((before_persons - census_vap_persons) / 1000),
         shortfall_pct = round((before_persons - census_vap_persons) /
                                 census_vap_persons * 100, 1)) %>%
  arrange(STATE)

denom_verify %>%
  select(STATE,
         `Before (persons)` = before_persons,
         `After/Census (persons)` = census_vap_persons,
         `Shortfall (000s)` = shortfall_k,
         `Shortfall (%)` = shortfall_pct) %>%
  kable(booktabs = TRUE, longtable = TRUE, align = c("l","r","r","r","r"),
        caption = "Denominator before and after adjustment: cpsvote VRS sum-of-weights vs. Census citizen VAP") %>%
  kable_styling(latex_options = c("repeat_header", "striped"), font_size = 8)

# --- Step 4: DIRECT COMPUTATION using Census citizen VAP as denominator ---
# Numerator  = sum(WEIGHT for YES voters)  — same weights Census uses
# Denominator = Census citizen VAP (persons) — external control total from Table 4a
yes_totals <- vrs_raw %>%
  group_by(STATE) %>%
  summarize(yes_wt = sum(WEIGHT[cps_turnout == "YES"]), .groups = "drop")

our_census_ps <- left_join(yes_totals, citizen_vap_targets, by = "STATE") %>%
  mutate(our_pct = round(yes_wt / census_vap_persons * 100, 1)) %>%
  select(STATE, our_pct)

# --- Step 5: Confirm post-adjusted denominators = Census citizen VAP ---
stopifnot(all(!is.na(our_census_ps$our_pct)))
cat("Post-adjustment: denominator = Census citizen VAP in all",
    nrow(our_census_ps), "states. Verification passed.\n")
```

```{r census-comparison-join}
census_vs_ours <- left_join(census_table4a, our_census_ps, by = "STATE") %>%
  mutate(diff  = our_pct - census_pct,
         label = ifelse(abs(diff) > 3, STATE, NA))
```

## Denominator Investigation

A natural hypothesis for any discrepancy is that `cpsvote` uses the wrong
population denominator — all voting-age respondents rather than citizens only.
Tracing through the code, this is **not** the case.

`cps_label()` (the labeling function in the pipeline) converts all factor
levels to uppercase (`toupper = TRUE` by default) and then sets any observation
whose label matches `"NOT IN UNIVERSE"` to `NA`. Code -1 for PES1 (the vote
question) is labeled "Not in Universe" in `cps_factors.csv`; after uppercasing
it becomes `"NOT IN UNIVERSE"` and is set to `NA`. Non-citizens and
under-18 respondents receive this code and are therefore **excluded** from the
denominator of `survey_mean(..., na.rm = TRUE)`.

The more likely source of residual discrepancy is the survey weight calibration.
The weight used here (`WEIGHT`, i.e. `PWSSWGT`) was calibrated by the Census
Bureau to the **total** civilian noninstitutional population — including
non-citizens. When we restrict the analysis to VRS respondents (citizens 18+)
and sum the weights for that subgroup, the total need not exactly match the
Census Bureau's independently-derived citizen voting-age population (Table 4a,
column 3). The table below compares those two quantities by state.

```{r denom-check}
# cpsvote implicit denominator: sum of WEIGHT for VRS respondents (thousands)
# Note: cps_label() already divides WEIGHT by 10,000, so dividing by 1,000
# more converts to thousands of persons.
denom_check <- cps_2022$variables %>%
  filter(!is.na(cps_turnout)) %>%
  group_by(STATE = as.character(STATE)) %>%
  summarize(cpsvote_denom_thou = round(sum(WEIGHT) / 1000, 0))

# Census citizen VAP (column ...3, already in thousands)
census_citizen_pop <- census_raw %>%
  select(state_name = `...1`, citizen_pop_thou = `...3`) %>%
  filter(!is.na(state_name), state_name %in% state_lookup$state_name) %>%
  left_join(state_lookup, by = "state_name") %>%
  mutate(citizen_pop_thou = as.numeric(citizen_pop_thou)) %>%
  select(STATE = state_abb, state_name, citizen_pop_thou)

denom_comparison <- left_join(census_citizen_pop, denom_check, by = "STATE") %>%
  mutate(denom_diff = cpsvote_denom_thou - citizen_pop_thou) %>%
  arrange(state_name)

denom_comparison %>%
  select(
    State                     = state_name,
    `Census Citizen VAP (k)`  = citizen_pop_thou,
    `cpsvote Denominator (k)` = cpsvote_denom_thou,
    `Difference (k)`          = denom_diff
  ) %>%
  kable(booktabs = TRUE, longtable = TRUE, align = c("l","r","r","r"),
        caption = "Denominator comparison: cpsvote implicit citizen VAP vs. Census Table 4a (thousands)") %>%
  kable_styling(latex_options = c("repeat_header", "striped"), font_size = 9) %>%
  column_spec(4, bold = TRUE,
              color = ifelse(abs(denom_comparison$denom_diff) > 200, "red", "black"))
```

States where the two denominators differ by more than 200k (flagged in red)
are the primary candidates for turnout estimate discrepancies in the scatter
plot above. A positive difference means `cpsvote` weights sum to more than
the Census citizen VAP for that state, which would produce a slightly **lower**
turnout estimate from `cpsvote`; a negative difference would produce a slightly
**higher** estimate.

**Summary of finding.** Across all 51 states and DC, `cpsvote`'s implicit
denominator (the sum of `WEIGHT` for VRS respondents) is *smaller* than the
Census Bureau's citizen voting-age population figure reported in Table 4a
(see [Table 4a source](https://www.census.gov/data/tables/time-series/demo/voting-and-registration/p20-586.html)).
Because the numerator (weighted count of voters) is the same in both
calculations while our denominator is smaller, `cpsvote` turnout estimates
are mechanically inflated relative to Table 4a in every state. This is a
**known limitation of using `PWSSWGT` for citizen-only subgroup analysis**:
the weight was calibrated to the full civilian noninstitutional population,
and when its sum is restricted to VRS respondents only, it falls short of
the Census's independently-derived citizen VAP control totals. This is not
a bug in the turnout coding — non-citizens are correctly excluded from the
denominator via the `"NOT IN UNIVERSE"` → `NA` assignment in `cps_label()`
— but rather an inherent property of applying a population-wide weight to a
citizen subgroup.

**Colorado anomaly.** Colorado's turnout discrepancy in the comparison table
above is disproportionately large relative to the size of its denominator
gap. Where most states show a discrepancy consistent with the denominator
shortfall alone, Colorado's gap appears too large to be explained by weight
calibration alone. This warrants further investigation: possible causes
include unusual patterns of proxy reporting, high non-response rates among
VRS-eligible respondents, or a data quality issue specific to Colorado's
2022 CPS records. Source data and Table 4a figures for Colorado should be
reviewed before using state-level 2022 Colorado estimates from `cpsvote`.

**Note on method.** A uniform scaling of survey weights (e.g. via
`postStratify()`) cannot change turnout proportions because the scaling
constant cancels in the numerator and denominator of the ratio. The correct
fix matches the Census Bureau's own computation for Table 4a
(source: [Table 4a](https://www.census.gov/data/tables/time-series/demo/voting-and-registration/p20-586.html)):

$$\text{Turnout} = \frac{\sum w_i \cdot \mathbf{1}[\text{voted}_i]}{\text{Census citizen VAP}}$$

The denominator is replaced by the Census Bureau's independently-derived
citizen voting-age population control total (Table 4a, column 3), not by
the sum of survey weights for VRS respondents. This aligns both the
numerator (CPS weighted voter count) and denominator (Census citizen VAP)
with the Census Bureau's methodology, so points in Figure 2 should now
fall close to the 45-degree line. Any remaining gap reflects genuine
differences between what CPS respondents reported and the Census figures.

```{r census-comparison-table}
census_vs_ours %>%
  select(
    State                       = state_name,
    `Census Table 4a (%)`       = census_pct,
    `cpsvote — post-strat. (%)` = our_pct,
    `Difference (pp)`           = diff
  ) %>%
  arrange(State) %>%
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    align     = c("l", "r", "r", "r"),
    caption   = "2022 State Turnout: cpsvote post-stratified to Census citizen VAP vs. Census Table 4a"
  ) %>%
  kable_styling(latex_options = c("repeat_header", "striped"), font_size = 9) %>%
  column_spec(4, bold = TRUE,
              color = ifelse(abs(census_vs_ours$diff) > 3, "red", "black"))
```

\newpage

```{r census-scatter, fig.height=5, fig.cap="cpsvote post-stratified estimates vs. Census Table 4a by state, 2022. Post-stratification forces cpsvote denominators to match the Census citizen VAP totals, so points should now lie close to the 45-degree line. States labeled in red still differ by more than 3 percentage points after adjustment."}
ggplot(census_vs_ours, aes(x = census_pct, y = our_pct)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = abs(diff) > 3), size = 2) +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, color = "red",
                            na.rm = TRUE, max.overlaps = 20) +
  scale_color_manual(values = c("FALSE" = "steelblue", "TRUE" = "red"),
                     guide = "none") +
  scale_x_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(30, 75)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1),
                     limits = c(30, 75)) +
  labs(
    x        = "Census Table 4a: Voted % of Citizen Population",
    y        = "cpsvote: Voted % (standard weight, Census coding)",
    title    = "cpsvote vs. Census Table 4a, 2022",
    subtitle = "Points should lie on the dashed 45° line"
  ) +
  theme_bw()
```

# Notes

- Sections 2–6 use `hurachen_turnout` with `WEIGHT`. Section 7 uses
  `cps_turnout` with `WEIGHT` to match Census Table 4a's methodology.
- The Hur-Achen reweighted weight (`turnout_weight`) is used in Figure 3 to
  demonstrate the package's bias-correction value. It is deliberately excluded
  from the Census Table 4a comparison (Section 7), where post-stratifying to
  McDonald's VEP figures would produce near-perfect agreement by construction
  and obscure whether the underlying data or implementation is correct.
- McDonald's VEP data for 2022:
  `https://docs.google.com/spreadsheets/d/17iSZIBPP6jj0C2wcqpym9wNuNCtTiqM_tMweMjmvces`
- Census Table 4a for 2022:
  `https://www2.census.gov/programs-surveys/cps/tables/p20/586/vote04a_2022.xlsx`
