---
title: "Voting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{voting}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  messages = FALSE, warnings = FALSE,
  fig.width = 6, fig.height = 3
)
library(knitr)
library(kableExtra)

```

```{r packages, include = FALSE, echo = FALSE}
library(cpsvote)
library(dplyr)
library(ggplot2)
library(here)
library(tidyverse)

```

# Using the CPS Voting and Registration Supplement to Examine Voting

The Current Population Survey's Voting and Registration Supplement, collected in November of every
federal election year, is an important source to understand many aspects of the American elections system: voter registration, 
voting turnout, modes of voting (Election Day, Early In Person, Voting By Mail), and barriers to registration and turnout. Because of its 
relative stability over time, large sample sizes, and quality of administration, the CPS is a uniquely valuable data resource to 
estimate voting turnout in the United States 
and in individual states, regions, and among key demographic groups. 

This vignette focuses on how the `cpsvote` package provides a 
simple avenue to examine voting turnout and mode of voting. The vignette
documents key data transformation decisions that are made so as to 
allow comparisons over years. 

# Overview

There are three challenges to use the CPS VRS for estimating voting turnout and mode of voting.

The first is an idiosyncratic coding method that the Census has used to code voting turnout that needs to be
understood by the user, or else basic descriptive statistics will not match those reported by the Census in
its documentation and reports. 

Second and related is a more complicated methodological issue that involves changes in the rate of non-response (as not voting) and resulting biases in CPS's turnout estimates which have grown over time. 
The appropriate Census recodes and statistical adjustments for non-response bias are automated as part of `cpsvote`.  A new survey weight is computed, 
`turnout_weight` that adjusts for these biases.

Third, CPS changed the way that it asked about _mode of voting_, using
two questions from 1996-2002, and a single question from 2004-present. `cpsvote` creates a consolidated variable, `VRS_VOTEMETHOD_CON`, that
codes for Election Day, Early In Person, and Voting By Mail from 1996-2018.

# Voting Turnout Using The CPS

The CPS has long used an "idiosyncratic" coding rule for reporting turnout, derived from the answer
to question "PES1" since 1996 (in 1994, the question was labeled "PES3").  The coding rule
is not clear from the CPS documentation, and without making the appropriate data transformations, any turnout estimates that are produced
will not match those in official Census communications. 

In the CPS codebook, the variable is reported in this way (from the 2016 documentation): 

In any election, some people are not able to vote because they are sick or busy or have some other reason, and others do not want to vote. Did (you/name) vote in the election held on Tuesday, November X, XXXX?

<table>
<tr>
<td>NAME</td> <td>   SIZE </td> <td>  DESCRIPTION </td>  <td>  LOCATION </td>
</tr>
<tr>
<td> PES1 </td> <td>   2  </td> <td> In any election, some people are  not able to vote because they are 
sick or busy or have some other reason, and others do not want to vote. Did (you/name) 
vote in the election held on Tuesday, November X, XXXX?} </td> <td> 951-952 </td>
<tr> <td> <td> <td> EDITED UNIVERSE: PRTAGE >=18 and PRCITSHP = 1, 2, 3, or 4<td> </tr>
<tr> <td> <td> <td> <td> </tr>
<tr> <td> <td> <td>VALID ENTRIES: <td> </tr> 
<tr> <td> <td> <td> 1  Yes <td> </tr>
<tr> <td> <td> <td> 2  No  <td> </tr>
<tr> <td> <td> <td> -1 Not in Universe <td> </tr> 
<tr> <td> <td> <td> -2 Don't Know <td> </tr>
<tr> <td> <td> <td> -3 Refused <td> </tr>
<tr> <td> <td> <td> -9 No Response <td> </tr> 
</table>

The key decision made by the Census staff has been to treat three of 
these non-response categories as "not voted", essentially adding these
cases to the denominator for a turnout estimate. 

The rule was carefully documented by two scholars, Aram Hur and Christopher Achen, in a 2013 article titled ["Coding Voter Turnout Responses in the Current Population Survey"](https://academic.oup.com/poq/article/77/4/985/1843466/).

Because some users of the `cpsvote` package may not be able to access this article, we reproduce the critical section from Hur and Achen (2013) that describes the coding choices below:

>In its official reports, however, the CPS does not follow the conventional academic coding rules for turnout responses. Instead, it treats Don’t Know, Refused, and No Response as indicating that the respondent did not vote. ... The Census Bureau’s decision to count the No Response individuals as nonvoters is consequential. No Response alone makes up 11.2 percent of the 2008 sample. Coding all of them, plus Don’t Know and Refused, as nonvoters reduces the estimated turnout rate by nearly 10 percentage points, cancelling most of the 12-point overreport in the original data.

**The `cpsvote` package provides an option to apply the Census coding scheme to creat two new columns, `cps_turnout` and `hurachen_turnout`.**

## How To Correct For Non-Response Bias in CPS Turnout Estimates

A related problem with the CPS turnout estimate, documented carefully by Professor Michael McDonald in a 2014 working paper^[Michael McDonald, 2014, "What's Wrong with the CPS?", Paper presented at the Annual Meeting of the American Political Science Association.] and at the [http://www.electproject.org/home/voter-turnout/cps-methodology](United States Elections Project's CPS Over-Report and Non-Response Bias Correction) page is that, over time, two biases have crept into the CPS, one from increasing non-response rates, the second from over-reports of turnout. 

From the Hur and Achen (2013) abstract:

>"The Voting and Registration Supplement to the Current Population Survey (CPS) employs a large sample size and has a very high response rate, and thus is often regarded as the gold standard among turnout surveys. In 2008, however, the CPS inaccurately estimated that presidential turnout had undergone a small decrease from 2004. We show that growing nonresponse plus a long-standing but idiosyncratic Census coding decision was responsible. We suggest that to cope with nonresponse and overreporting, users of the Voting Supplement sample should weight it to reflect actual state vote counts."

Hur and Achen suggest a complex post-stratification adjustment to the data that will adjust for these biases:

>We recommend dropping all categories of missing turnout response, and then poststratifying the remaining CPS sample so that the survey turnout rate in each state matches the corresponding state VEP turnout.

Professor Michael McDonald of the University of Florida helpfully provides guidance on this more complex procedure. He provides <Commentary, Guidelines, and Stata Code>[link](http://www.electproject.org/home/voter-turnout/cps-methodology) at his website. 

**The Hur and Achen corrections have been integrated into the `cpsvote` package.**

## Recoding and Reweighting Voting Turnout Using `cpsvote`

`cpsvote` provides two commands that will create two new turnout 
variables (`cps_turnot` and `hurachen_turnout`) and produce a Hur-Achen weight (`turnout_weight`). 

### A One-Step Guide for People Who Don't Read Documentation

1. `cps <- cps_load_basic()` Will load all available years of data (and if not available, will download) and perform a set of additional data labeling and transformations to prepare the CPS VRS data from 1994 - 2018 for analysis. 
2. `cps <- cps_load_basic(years = seq(2000, 2008, 4))` will load the 2000, 2004, and 2008 CPS.

### Recoding Turnout: Detailed Description

We have created a function that will produce two new
columns corresponding to the "CPS" method of coding turnout
and the "Hur-Achen" alternative.

```
cps_recode_vote(
  data,
  vote_col = "VRS_VOTE",
  items = c("DON'T KNOW", "REFUSED", "NO RESPONSE")
)
```

#### Arguments

1. _data_: the input data set
2. _vote_col_	which column contains the voting variable
3. _items_	which items should be "No" in the CPS coding and NA in the Hur & Achen coding

#### Output

A data frame data with two columns attached, `cps_turnout` and `hurachen_turnout`, voting variables recoded according to the process above. If the command is issued without any parameters beyond the data frame,
the default column names are assumed. 

### Reweighting the CPS To Adjust For Non-Response Bias

We have created a function that will produce a new
survey weight, following the suggested method by Hur 
and Achen.

`cps_reweight_turnout(data)`

#### Arguments

1. _data_: the input data set, which must contain columns named `YEAR`, `STATE`, and `hurachen_turnout`

#### Output

A data frame data with one column attached, `turnout_weight` which can be used in lieu of the 
Census provided weight (name `WEIGHT` after renaming by `cps_vote`).

## Sample Usage: Estimating Voter Turnout in 2016

Start by reading the files and performing specialized `cpsvote` recoding, create new survey weights. 

```{r setup2, echo = TRUE}
#library(cps)
cpsvote::cps_download_data(years = 2016) # Download the 2016 CPS, if not already downloaded
cps2016 <- cpsvote::cps_read(years = 2016) # Read in CPS 2016. Ignore the warning since we don't care
                                           # about joining any files
## Apply the turnout recodes
cps2016 <- cps_recode_vote(cps2016)

## Create the new survey weight
cps2016 <- cps_reweight_turnout(cps2016)

## Create factors out of categorical variables 
cps2016 <- cps_label(cps2016)

## Refactor questions that differ slightly across years
cps2016 <- cps_refactor(cps2016)

## NOTE a one-step method for reading in the
##  data is:
##
## cps2016 <- cps_load_basic(years = 2016)

## Dichotomous turnout variables for calculating means
cps2016 <- cps2016 %>%
  mutate(cps_turnout_d = case_when(
    cps_turnout == "YES" ~ 1,
    cps_turnout == "NO" ~0),
    hurachen_turnout_d = case_when(
    hurachen_turnout == "YES" ~ 1,
    hurachen_turnout == "NO" ~ 0)
  )

## SURVEY WEIGHTS
# The CPS data MUST be weighted. The "weighting" vignette provides additional information 
                # about using using the survey or srvyr packages to produce valid statistical results

library(survey)
library(srvyr)  

cps2016_surveyCW <-                                             # SURVEY Package Syntax
    svydesign(                                                  # Census weight
        ~ 1 , 
        data = cps2016, 
        weights = ~ WEIGHT
    )

cps2016_surveyHW <-                                             # SURVEY Package Syntax
    svydesign(                                                  # Hur Achen weight
        ~ 1 , 
        data = cps2016, 
        weights = ~ turnout_weight
    )

cps2016_srvyrHW <- as_survey_design(cps2016, weights = turnout_weight)  # SRVYR Package Syntax, 
                                                                        # Hur and Achen weight
```

_Generate unweighted, CPS weighted, and Hur/Achen weighted estimates of voter turnout._ 

```{r turnout, echo = TRUE}

turnout_estimates <- cps2016 %>%
  summarize(Unweighted = mean(cps_turnout_d, na.rm = TRUE),
            Census_Weighted = svymean(cps_turnout_d, cps2016_surveyCW, na.rm = TRUE),
            Hur_Achen_Weighted = svymean(hurachen_turnout_d, cps2016_surveyHW, na.rm = TRUE))

kable(turnout_estimates)
```

The table shows the slight overestimate of turnout using the Census
method, because it fails to account for growing non-response bias. 

_Generate Hur/Achen weighted turnout estimates by racial groups_

```{r turnout_race, echo = TRUE}

cps2016_srvyrHW %>%
  filter(RACE %in% c("WHITE", "BLACK", "AMERICAN INDIAN OR ALASKA NATIVE",
                     "ASIAN, PACIFIC ISLANDER, OR NATIVE HAWAIIAN")) %>%
  group_by(RACE) %>%
  summarize(turnout = survey_mean(hurachen_turnout_d, na.rm = TRUE)) %>%
  ggplot(., aes(x = RACE, y = turnout)) +
  geom_bar(stat = "identity") + 
  scale_x_discrete(labels = c("WHITE"= "White",
                                "BLACK" = "Black", 
                                "AMERICAN INDIAN OR ALASKA NATIVE" = "Amer Indian/Alakan",
                                "ASIAN, PACIFIC ISLANDER, OR NATIVE HAWAIIAN" = "Asian/Hawaii/PacIsland")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "", y = "Turnout", title = "Turnout by Race, 2016")


```

_A simple map of voter turnout by state._ 

```{r turnout_statemap, echo = TRUE}
library(usmap)

cps2016_srvyrHW %>%
  mutate(state = STATE) %>%
  group_by(state) %>%
  summarize(turnout = survey_mean(hurachen_turnout_d, na.rm = TRUE)) %>%
  plot_usmap(data = ., values = "turnout", color = "blue") +
  scale_fill_continuous(low = "white", high = "blue", name = "Turnout Rate") +
  theme(legend.position = "right") + labs(title = "Voting Turnout in the United States, 2016")

```


# Voting Mode in CPS: Election Day, Early In Person, and Absentee/Voting by Mail

Capturing the difference between casting an in-person ballot on Election Day, an in-person ballot early at a designated location, and a ballot cast "by mail" (either absentee or 
in full vote by mail states) is difficult given the various ways that states allow for
a ballot to be cast. For instance, in a state that allows you to appear at a county 
office before or on Election Day, request an absentee ballot, and complete and submit it 
at that time, is that voting "in person" or "by mail"?

We will put aside these complexities to discuss how the CPS has captured this voting activity, what is called the "mode of voting", over time. There is one major change to
pay attention to.

From 1996-2002, the CPS asked a single question (PES4), that included the options "In person on election day," "In person before election day", and "Voted by mail (absentee)."

Starting in 2004 and continuing through the present, the CPS started to ask about the _time_ that a respondent cast their ballot (PES5: "On Election Day", "Before Election Day") and the _method_ by which the ballot was cast (PES6: "In Person", "By Mail"). 

From 1996-2002, `cpsvote` uses the values in PES4. From 2004 to the present, the
answers to PES5 and PES6 are combined to create a three category vote mode variable ("Election Day", "Early", "By Mail"). The package assumes that an answer of "By Mail" counts as absentee / voting by mail, regardless of "when" the respondent said the ballot was cast.

The table below provides an overview of these coding decisions:

<table>
<caption> A Unified Mode of Voting Variable for CPS </caption>
<tr>
<td> <b> `cpsvote` Output <td colspan = "3"> <b> Census Input </tr>
<td>  <td>  <b> 2004-present <td> <td> <b> 1996 - 2002  </b> </tr>
<tr> </tr>
<tr>
<td>  <b>VRS_VOTEMETHOD_CON      <td> PES5     <td> PES6    <td> PES4  </b> </tr>
<tr>
<td> <b>Election Day</b>       <td> In Person <td> Election day <td> In person on election day</tr>
<tr>
 <td> <b>Early </b>              <td> In Person <td> Before Election Day <td> In person before election day <tr> 
 <td> <b>By Mail </b>            <td> By Mail  <td> Election Day, Before Election Day <td> Voted by mail (absentee)</tr>
</table>

<!-- \begingroup
\setlength{\tabcolsep}{10pt} % Default value: 6pt
\renewcommand{\arraystretch}{3.7} % Default value: 1
\begin{table}[]
\begin{tabular}{lcccccc}
                                                          & \multicolumn{1}{l}{}                 & \multicolumn{2}{c}{PES5}                                    & \multicolumn{1}{l}{} & \multicolumn{2}{c}{PES6}                                                   \\ \cline{3-4} \cline{6-7} 
                                                          & \multicolumn{1}{l}{}                 & \multicolumn{1}{l}{In-person} & \multicolumn{1}{l}{By-mail} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{Election day} & \multicolumn{1}{l}{Before election day} \\ \cline{3-4} \cline{6-7} 
\multicolumn{1}{l|}{\multirow{3}{*}{\rotatebox[origin=c]{90}{VRS\_VOTEMETHOD\_CON}}} & \multicolumn{1}{c|}{Election day}    & $\Checkmark$                  &                             &                      & $\Checkmark$                     &                                         \\
\multicolumn{1}{l|}{}                                     & \multicolumn{1}{c|}{Early in-person} & $\Checkmark$                  &                             &                      &                                  & $\Checkmark$                            \\
\multicolumn{1}{l|}{}                                     & \multicolumn{1}{c|}{Vote-by-mail}    &                               & $\Checkmark$                &                      & $\Checkmark$                     & $\Checkmark$                           
\end{tabular}
\end{table}
--> 



```{r earlyvoting_data_load}
cps10k <- cps_sample_10k

# Create region codes following Census guidelines

cps10k <-  cps10k %>% 
  mutate(census_region = case_when(
    STATE %in% c("ME", "NH", "VT", "MA", "CT", "RI", 
                   "NY", "PA", "NJ") ~ "Northeast",
    STATE %in% c("ME", "DE", "WV", "DC", "VA", "NC", "SC", "GA", "FL",
                   "KY", "TN", "MS", "AL", 
                   "OK", "AR", "LA", "TX") ~ "South",
    STATE %in% c("WI", "MI", "IL", "IN", "OH", 
                   "ND", "MN", "SD", "IA", "NE", "MO", "KS") ~ "Midwest",
    STATE %in% c("MT", "ID", "WY", "NV", "UT", "CO", "AZ", "NM", 
                   "WA", "OR", "CA", "AK", "HI") ~ "West"),
    census_division = case_when(
      STATE %in% c("ME", "NH", "VT", "MA", "CT", "RI") ~ "New England", 
      STATE %in% c("NY", "PA", "NJ") ~ "Middle Atlantic",
      STATE %in% c("ME", "DE", "WV", "DC", "VA", "NC", "SC", "GA", "FL") ~ "South Atlantic",
      STATE %in% c("KY", "TN", "MS", "AL") ~ "East South Central", 
      STATE %in% c("OK", "AR", "LA", "TX") ~ "West South Central",
      STATE %in% c("WI", "MI", "IL", "IN", "OH") ~ "East North Central", 
      STATE %in% c("ND", "MN", "SD", "IA", "NE", "MO", "KS") ~ "West North Central",
      STATE %in% c("MT", "ID", "WY", "NV", "UT", "CO", "AZ", "NM") ~ "Mountain", 
      STATE %in% c("WA", "OR", "CA", "AK", "HI") ~ "Pacific")
  )

    
    # Set up the survey design. Weights must be used for the CPS
#
cps10k_weight <- as_survey_design(cps10k, weights = WEIGHT)

```

```{r modesbyyear_line}

# Graph 1: Rate of All Modes of Voting By Year (line graph)

cps10k_weight %>%
  filter(YEAR > 1994 & !is.na(VRS_VOTEMETHOD_CON)) %>%
  group_by(YEAR, VRS_VOTEMETHOD_CON) %>%
  summarize(value = survey_mean(na.rm = TRUE)) %>%
  ggplot(aes(x = YEAR, y = value, col = VRS_VOTEMETHOD_CON, group = VRS_VOTEMETHOD_CON)) +
  geom_line(size = 1.5) +
  geom_point(aes(x = YEAR, y = value, color = VRS_VOTEMETHOD_CON), size = 2) +
  scale_x_continuous(breaks = seq(1996, 2018, by = 2)) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "The Growth of Early Voting, 1996 - 2018", 
       subtitle = "Source: Current Population Survey, Voting and Registration Supplement",
       color = "Mode of Voting",
       y = "",
       x = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, family = "Times", face = "bold.italic", colour = "red"),
        legend.position = c(.2,.5), 
        legend.background = element_rect(), 
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10)) 
```

```{r voteathome, eval = FALSE}
cps10k_weight %>%
  filter(YEAR > 1994 &  !is.na(VRS_VOTEMETHOD_CON)) %>%
  group_by(YEAR, census_region, VRS_VOTEMETHOD_CON) %>%
  summarize(value = survey_mean(na.rm = TRUE)) %>%
  filter(VRS_VOTEMETHOD_CON == "By Mail") %>%
  ggplot(aes(x = YEAR, y = value, col = census_region, group = census_region)) +
  geom_line(size = 1.5) +
  geom_point(aes(x = YEAR, y = value, color = census_region), size = 2) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1996, 2018, by = 2)) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Regional Use of Vote At Home, 1996 - 2018", 
       subtitle = "Source: Current Population Survey, Voting and Registration Supplement",
       color = "Region") +
  theme(plot.title = element_text(size = 20, family = "Times", face = "bold.italic", colour = "red"),
        legend.position = c(.1,.8), legend.background = element_rect(),  
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10)) +
  ylab("") + xlab("")  

```

## Canyon Maps can go here


```{r}
library(albersusa) 
library(sf)
library(ggthemes)
map_data <- cps2016_srvyrHW %>%
  filter(YEAR == 2016 & !is.na(VRS_VOTEMETHOD_CON)) %>%
  group_by(STATE, VRS_VOTEMETHOD_CON) %>%
  summarize(value = survey_mean(na.rm = TRUE)) %>%
  merge(albersusa::usa_sf("laea"), by.x = "STATE", by.y = "iso_3166_2")


sf::st_as_sf(map_data) %>% 
  filter(VRS_VOTEMETHOD_CON == "BY MAIL") %>%
  ggplot(aes(fill = value)) +
  geom_sf() +
  ggthemes::theme_map() + 
  theme(legend.position = "top",
        text = element_text("times")) + 
  scale_fill_gradient(low = "sky blue", high = "navy", na.value = NA, name = "Vote-by-Mail Usage", labels = scales::percent)

  
  
sf::st_as_sf(map_data) %>% 
  filter(VRS_VOTEMETHOD_CON == "EARLY") %>%
  ggplot(aes(fill = value)) +
  geom_sf() +
  ggthemes::theme_map() + 
  theme(legend.position = "top",
        text = element_text("times")) +
  scale_fill_gradient(low = "yellow", high = "red", na.value = NA, name = "Early Voting Usage", labels = scales::percent)

  
sf::st_as_sf(map_data) %>% 
  filter(VRS_VOTEMETHOD_CON == "ELECTION DAY") %>%
  ggplot(aes(fill = value)) +
  geom_sf() +
  ggthemes::theme_map() + 
  theme(legend.position = "top",
        text = element_text("times")) +
  scale_fill_gradient(low = "lavender", high = "purple", na.value = NA, name = "Election Day Voting Usage", labels = scales::percent)

```

